\bf The usage of bootstrap method in shape-restricted regression

### 2023-08-14 Manuscript/manuscript.tex
Update manuscript.tex

It has been proved that the bootstrap method can be used as a good estimator for the pointwise confidence interval in linear regressions and non-linear regressions\cite{efron1979bootstrap}\cite{davidson1999bootstrap}, and some examples in certain configurations are given.\cite{ruhe2019bootstrap}\cite{ma2019inference}\cite{dugas2010pointwise} But in certain special cases(i.e. shape-restricted splines regressions) there is no certain resource have not prove that it is a good method to build pointwise confidence interval via bootstrap.

Shape-restricted splines regressions are piecewise polynominal, differentiable to a certain degree and have certain restrictions like monotonicity or convexity.\cite{meyer2008inference} Given these conditions, the application of the bootstrap method to estimate pointwise confidence intervals for specific time points within the regression cannot be guaranteed to be suitable. And in the rest of the article we gives out an example of shape-restricted regression and the method to fit confidence interval for certain time points in the function.

@@

In some situations the regression is forced to be shape-restricted to improve the predictive performance and reduce overfitting if the underlying regression function takes the specific form. Some regressions are naturally constrained to convexity or concativity in a lot of disciplines like economics, biology, and psychological studies, and other ares.\cite{jieying2022heteroscedastic} \cite{guntuboyina2018nonparametric} Some popular examples like cost function and profit function in economics\cite{gallant1984imposing}, precipitation data analysis\cite{molitor2002bayesian}. In this article we gives out an example of the relationship between income and the age of Canadian workers. The data is given from R package SemiPar. For those situations, Regression splines are smooth, flexible, and parsimonious nonparametric function estimators, and it offers an effective approach to enhance the inference of shape-restricted regression. \cite{meyer2008inference} In this dataset, a nondecreasing relationship between the age of workers and the expected logarithm of the income was considered. \cite{wang2021shape}  Splines2 provides implementation of those spline basis functions via R language, and we use that to build a non-decreasing curve by i-spline that force the coefficients to be non-negative with degree of freedom = 6. Then, we selected 10 time points, denoted as $t$, with equal intervals from the dataset, corresponding to one of them is the regression time points \(f'(t)\) that fitted by iSpline. We also performed bootstrap resampling 1000 times to obtain the 95\% quantile pointwise confidence interval for each points. In the figure below, the red line represents the fitted i-spline regression, the yellow line is the smooth regression, and we build ten pointwise confidence intervals for 10 certain time points. The result is shown in Figure ~\ref{fig:semipar}.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.8\textwidth]{DataAndRegressions.png}
  \caption{The generated dataset, true functions and fitted i-splines regressions: the shaded green curve represents the true functions, the pink points are the data pair \((x, f'(x))\) where \(f'(x)\) is generated by \(f(x)\) add noise, and the red curves are the fitted i-splines regression curve with degrees of freedom\(= 6,10,15\), from top to bottom}
  \label{fig:DataAndRegressions}
\end{figure}

@@

Our findings support the utility of the bootstrap method to build the pointwise confidence intervals in cases where regression models have specific constraints such as monotonicity or curvature. However, it's important to consider the assumptions and limitations of the bootstrap method and validate its performance for specific datasets and applications.

>>>>

It has been proved that the bootstrap method can be used as a good estimator for
the pointwise confidence interval in linear regressions and non-linear regressions
\cite{efron1979bootstrap}\cite{davidson1999bootstrap}, and some examples in certain
configurations are given.\cite{ruhe2019bootstrap}\cite{ma2019inference}
\cite{dugas2010pointwise} But in certain special cases(i.e. shape-restricted splines 
regressions) there is no certain resource have not prove that it is a good method to
build pointwise confidence interval via bootstrap.

Shape-restricted splines regressions are piecewise polynominal, differentiable to a
certain degree and have certain restrictions like monotonicity or convexity.
\cite{meyer2008inference} Given these conditions, the application of the bootstrap
method to estimate pointwise confidence intervals for specific time points within 
the regression cannot be guaranteed to be suitable. And in the rest of the article
we gives out an example of shape-restricted regression and the method to fit confidence
interval for certain time points in the function.

@@

In some situations the regression is forced to be shape-restricted to improve
the predictive performance and reduce overfitting if the underlying regression 
function takes the specific form. Some regressions are naturally constrained to
convexity or concativity in a lot of disciplines like economics, biology, and 
psychological studies, and other ares.\cite{jieying2022heteroscedastic} 
\cite{guntuboyina2018nonparametric} Some popular examples like cost function and 
profit function in economics\cite{gallant1984imposing}, precipitation data analysis
\cite{molitor2002bayesian}. In this article we gives out an example of the 
relationship between income and the age of Canadian workers. The data is given from
R package SemiPar. For those situations, Regression splines are smooth, flexible, 
and parsimonious nonparametric function estimators, and it offers an effective approach
to enhance the inference of shape-restricted regression. \cite{meyer2008inference} 
In this dataset, a nondecreasing relationship between the age of workers and the 
expected logarithm of the income was considered. \cite{wang2021shape}  Splines2 provides
implementation of those spline basis functions via R language, and we use that to build 
a non-decreasing curve by i-spline that force the coefficients to be non-negative with
degree of freedom = 6. Then, we selected 10 time points, denoted as $t$, with equal 
intervals from the dataset, corresponding to one of them is the regression time points
\(f'(t)\) that fitted by iSpline. We also performed bootstrap resampling 1000 times to
obtain the 95\% quantile pointwise confidence interval for each points. In the figure
below, the red line represents the fitted i-spline regression, the yellow line is the
smooth regression, and we build ten pointwise confidence intervals for 10 certain time
points. The result is shown in Figure ~\ref{fig:semipar}.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.8\textwidth]{DataAndRegressions.png}
  \caption{The generated dataset, true functions and fitted i-splines regressions
  : the shaded green curve represents the true functions, the pink points are the
  data pair \((x, f'(x))\) where \(f'(x)\) is generated by \(f(x)\) add noise,
  and the red curves are the fitted i-splines regression curve with degrees of 
  freedom\(= 6,10,15\), from top to bottom}
  \label{fig:DataAndRegressions}
\end{figure}

@@

Our findings support the utility of the bootstrap method to build
the pointwise confidence intervals in cases where regression models
have specific constraints such as monotonicity or curvature. However
, it's important to consider the assumptions and limitations of the
bootstrap method and validate its performance for specific datasets and applications.
